Species motif participation provides unique information about species risk of extinction

### 2023-06-06 manuscript/reorder_main_text.tex
Updates from Overleaf

% Now 286 words. Max 350. 
    \indent{1. Loss of species in food webs can set in motion a cascade of additional, secondary,  extinctions.A species' position in a food web (e.g., its trophic level or number of interactions) is known to affect its ability to persist following disturbance. Theses simple measures, however, offer only a coarse description of how species fit into their community. Therefore one would expect that more detailed structure measures, such as a species' participation in three-species motifs (meso-scale structures which provide information on a species' direct and indirect interactions) will also be related to its probability of persistence.\\
    2. Disturbances, such as species extinctions, affecting the basal resources have particularly strong effects on the rest of the food web. However, how the disturbance branch out and affect consumer persistence is depending on the structural pattern of species interactions in several steps. The magnitude, e.g. proportion of basal resources lost, will likely also affect the outcome.\\     
    3. Here, we analyse whether a consumer's risk of going secondary extinct after removal of basal resources depends on the consumer's motif participation and how this relationship varies with the severity of disturbance.\\
    4. We show that consumer species which participate more frequently in the direct competition motif and less frequently in the omnivory motif generally have higher probability of persistence following disturbance to basal resources.\\
    5. However, both the strength of the disturbance and the overall network structure (i.e., connectance) affect the strength and direction of relationships between motif participation and persistence.\\
    6. Motif participation therefore captures important trends in species persistence and provides a rich description of species' structural roles in their communities, but must be considered in the context of network structure as a whole and of the specific disturbance applied}

@@

    Motifs are unique patterns of $n$ interacting species that correspond to different sets of direct and indirect interactions, providing a holistic description of meso-scale network structure~\citep{Stouffer2007,Stouffer2012}.
    Some motifs are over-represented in stable food webs, suggesting that these motifs provides stability~\citep{Borrelli2015a}. This further suggest that the more frequently species are participating in these motifs, the more stable the food web would be.
    % Indeed, dynamic simulations show that participation in some motifs can provide greater persistence after the loss of a single species~\citep{Cirtwill2022Oikos}, suggesting that motif participation may be related to persistence after other types of disturbance.
    The vector of frequencies with which a species appears in each motif --its \emph{motif participation}-- describes the species' position in the meso-scale structure of a food web and, like degree and trophic level, is easily calculated once the structure of the network is known.
    Importantly, motif participation incorporates information on the positions of focal species' interaction partners, as well as potential indirect interactions between species.
    Two species might have the same trophic level and the same number of prey, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists (for an example see Fig.~\ref{fig:concept}).
    Since specialists are usually more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than will a predator of generalists.
    This extra information on the focal species' interaction partners may thus provide extra insight when predicting extinction risk.

    %There are 13 possible three-species motifs (the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}, 
    We define motif participation using the well-studied three-species motifs~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}. 
    Four of these motifs --the three-species chain, apparent competition, direct competition, and omnivory-- are especially interesting as they are by far the most common in empirical networks~\citep{Stouffer2007, giling2019plant} and more stable than other motifs (i.e., are more likely to retain all three species in the motif) when modelled in isolation~\citep{Borrelli2015a}.
    These four motifs also have clear ecological interpretations.
    The three-species chain indicates vertical flows of energy and biomass in a network and captures the possibility of indirect interactions across multiple trophic levels (e.g., between a plant and a predator of the plant's herbivore).
    The dynamics of three-species chains in isolation are well-understood from Lotka-Volterra models, and stable equilibria with all three species persisting can often be identified~\citep{Borrelli2015a,Hsu2015}.
    In general, food chains are more likely to be stable when predator-prey body-mass ratios are smaller~\citep{Jonsson1998}.
    The direct competition motif, where two consumers share the same resource, captures the possibility of indirect interactions between consumers by depletion of the shared resource in competitive exclusion~\citep{Hsu2015}. 
   ( Despite the potential for exclusion where one predator is competitively superior, this motif is highly robust to changes in interaction strength and therefore has high quasi-sign stability~\citep{Borrelli2015a}.)
    The apparent competition motif indicates a predator which consumes two resources and, by being somewhat generalist, may be able to persist following the loss or reduction of one prey.
    However, there may also be indirect effects between the linked prey that make the predator more likely to go extinct~\citep{Melian2002}.
    Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes  the potential for a plant to have both direct and indirect effects on the predator. 
    Omnivory is less stable in isolation than the other three motifs~\citep{Borrelli2015a}, though there is a region of stable equilibrium for Lotka-Volterra models of omnivory~\citep{Hsu2015}.
    In whole food webs, frequent omnivory has been proposed to both increase and decrease overall stability in different conditions~\citep{McCann1997,Emmerson2004,Monteiro2016}.
    As most of the above studies either model motifs in isolation or examine overall effects on the whole food web, it remains unclear how a particular species' motif participation will affect its specific extinction risk.

    Previous work using dynamical simulations has shown that participation in some motifs is associated with longer persistence after the loss of a single species in the food web~\citep{Cirtwill2022Oikos}, but the relationship between motif participation and persistence in the face of other disturbances remains unclear.
   As the basal resource community is the foundation upon which all other species depend, and loss of basal resource diversity typically causes declines in both abundance and richness of all types of consumers~\citep{scherber2010bottom,Dobson2009food, Mduma1999food, Georgiadis2007}.
    Moreover, disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant}, and consumer diversity \citep{scherber2010bottom, Baiser2016,li2020bottom}.
    It is therefore crucial to understand how bottom-up effects from disturbance to basal resources affect the rest of the food web.  Here, we are interested in understanding if motif participation can provide additional information on how extinction risk of consumer species are affected by the loss of basal resources, i.e., the effects of bottom-up disturbances.
    As well as identifying patterns of motif participation that are associated with ability to persist after a loss of basal resources, we are particularly interested in whether the strength of  disturbance affects relationships between motif participation and persistence.
    For example, participation in one motif might be associated with greater extinction risk after a mild disturbance but be associated with less risk when when more basal resources are lost if the direct and indirect interactions within the motif amplify the effect of small disturbances while providing `back-up' resources when many prey are removed by a strong disturbance.

@@

        The removal of cycles also affects which motifs can appear in the network. 
        Any motif which contains a cycle is impossible in an acyclic network.
        This means that only four motifs can appear: apparent competition, direct competition, omnivory, and the three-species chain.
        Fortunately, they are also the motifs that are most common in empirical networks (\citealp[]{Stouffer2007}, have been most studied in isolation (e.g.,~\citep{Hastings1991,Holt1987,Bascompte2005,Polis1989,Zabalo2012,Lefevre2009,Holt1997,Kondoh2008,McKinnon2013,Laws2013}), have the clearest ecological interpretations and, because of these three properties, are of greatest interest to us.

        \subsubsection*{Modelling consumers' risk of secondary extinction} [[trimmed down to here]]

@@

            We also confirmed previously-known relationships between persistence and other measures of network structure or a species' participation in the network.
            To identify the relationship between large-scale network structure and mean persistence within a network, and how the strength of disturbance might affect this relationship, we fit a GLM relating mean persistence to network size, connectance, disturbance, and their interactions.
            As the three-way interaction was significant, we did not simplify the model. 
            To confirm known relationships between persistence and in-degree or trophic level in our simulated networks and identify how they might vary with different strengths of disturbance, we fit two GLMs with binomial error distribution relating a consumer's persistence to in-degree or STL, disturbance, and the interaction between them (\emph{Appendix S3.1}, Eq. S2-3).
            As with the GLMs relating persistence to motif participation and motif profiles, we initially fit GLMMs including random effects of network size and connectance and network identity but removed these terms due to model singularity.

@@

        The average persistence of all consumers was lower in networks with higher connectance, did not vary significantly with network size (Table S4, \emph{Appendix S6.1}), and  varied with the frequencies of different motifs in this profile.
        Specifically, average consumer persistence was higher in networks with higher proportions of the apparent competition motif, lower in networks with higher proportions of the omnivory motif, and was not significantly related to the frequencies of the three-species chain or direct competition motifs (Fig.~\ref{fig:lm_CS}, Table S5, \emph{Appendix S6.2}). 
        There was no significant interaction between the proportion of any motif in the network profile and level of disturbance).

@@

    Disturbance strength is often neglected when predicting risk of extinction, but proved important in our results.
    Several of the relationships we demonstrated between probability of persistence and species roles changed direction when moving from weak to strong disturbance.
    These simulation results echo other warnings about `tipping points' where previously beneficial traits can become risky~\citep{Latty2019,Golubski2016,Tylianakis2014}.
    For example, at low levels of disturbance a high degree may be protective since species can compensate if one or a few prey are lost. 
    At high disturbance, however, a high-degree species is likely to be affected by disturbances through many paths and therefore have lower persistence.
    This context-dependence contrasts with the general assumption that species with more resources will be less vulnerable~\citep{Allesina2012,binzer2011susceptibility}.
    These disturbance effects mean that efforts to define a set of target species for conservation interactions must take the current (and future) level of disturbance the community experiences into account.

@@

    The authors thank Gy\"{o}rgy Barab\'{a}s for excellent discussion of the ideas in this manuscript. A{\AA} is supported by the Swedish Research Council for Sustainable Development Formas (\#2015-01262) granted to AE. ARC is supported by a Finnish Academy postdoctoral research grant (\#332999).

@@

        \begin{figure}[hb!]
        \centering
        \includegraphics[width=.9\textwidth]{figures/Figure_1.eps}
        \caption{Conceptual figure of a toy food web. \textbf{a)} This food web contains examples of the four motifs: apparent competition (orange), three-species chain (purple), omnivory (red), and direct competition (blue). These motifs capture different ways in which disturbances to basal resources can propagate through the network. For example, a disturbance to the bottom species in a three-species chain will have a direct effect on the middle species but only indirectly affect the top species while a disturbance to the bottom species in an omnivory motif will have a direct effect on the middle species and direct and indirect effects on the top species. 
        Note that the nodes participate in more motifs than the ones highlighted and that a species' motif participation vector is the count of all motifs of each type in which the species appears. \textbf{b)} In the same food web, species with identical degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only three chains while species B (dark grey node with its direct consumption links colored dark grey) participates in seven (see inserted table for counts of participation in all four motifs; species A has the motif participation vector [3, 3, 0, 1] while species B has the motif participation vector [3, 7, 0, 1]).
        Despite having equal degree and trophic level, species A may be at greater risk of going extinct than species B because of its dependence on specialist prey species. Two-thirds of the prey of species B are generalists and therefore more likely to survive a disturbance to basal resources. By including information about indirect interactions, motif profiles reflect this difference between species A and B while trophic level and degree do not.}
    \label{fig:concept}
    \end{figure}

@@

        \begin{figure}[hb!]
        \centering
        \includegraphics[width=\textwidth]{figures/persistence_motif_participation.eps}
        \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on probability of persistence following disturbance to basal resources (y-axis). The effect of participating in each motif is based on the Equation~\ref{propreq}. The different colored lines indicate the probability of extinction of basal species, from $\pi_{disturbed} = 0.1$ (top, purple; no disturbance) to $\pi_{disturbed} = 0.5$ (bottom, yellow; high disturbance). 95\% confidence intervals for each line are shown in lighter shading. Note that omnivory made up a smaller proportion of species' roles than other motifs; lines are plotted over the observed ranges of motif participation.}
    \label{fig:prop_lmer_all}
    \end{figure}

    \begin{figure}[hb!]
    \centering
        \includegraphics[width=\textwidth]{figures/Fig4.pdf}
        \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species. Columns show the result for different motifs, rows show results for low, medium, and high connectance, and colour indicates the strength of disturbance to basal resources ($\pi_{disturbed}$; probability of extinction for a basal resource). The dotted, vertical lines indicate zero on the x-axis. A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increases persistence.}
        \label{fig:density_prop}
    \end{figure}

    \begin{figure}[ht!]
        \centering
        \includegraphics[width=\textwidth]{figures/persistence_vs_SC_lm.eps}
        \caption{The probabilities of persistence for a particular consumer (\textbf{A, B}) and the average probability of persistence across all consumers (\textbf{C-E}) varied with other elements of network structure as well as with motif participation.
        \textbf{A} Persistence of consumers decreased with increasing STL at all levels of disturbance.
        \textbf{B} Persistence of consumers increased with increasing in-degree when the probability of extinction for basal resources was low but decreased with increasing in-degree when the probability of extinction was high.
        \textbf{C,D} Mean persistence of consumers in a network decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance. There was no significant relationship between mean persistence and network size.
        \textbf{E} Mean persistence of consumers decreased with increasing proportions of omnivory in the network's motif profile and increased with increasing proportions of other motifs. Persistence also decreased significantly with increasing probability of basal species extinction but there was no significant interaction; we therefore show trends only for $\pi$=0.1.}
        \label{fig:lm_CS}
    \end{figure}

    \begin{figure}[hb!]
        \centering
        \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
        \caption{Motif participation correlated with in-degree and trophic level, and both of these simpler role measures were also correlated with persistence. \textbf{A)} The proportion of omnivory increased with increasing degree while all other proportions decreased. \textbf{B)} The proportions of omnivory and direct competition decreased with increasing STL while the proportions of apparent competition and three-species chains increased. \textbf{C-D)} The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition, where the interaction was not significant). To cover the largest range of slopes, we show relationships over connectance for the smallest networks (\textbf{C}) and over network size for the most-connected networks (\textbf{D}). See also Fig. S6.}
        \label{fig:motifs_vs_TL_and_deg}
    \end{figure}

>>>>

% Now 281 words. Max 350. 
    1. Loss of species in food webs can set in motion a cascade of additional (secondary)  extinctions.
    2. A species' position in a food web (e.g., its trophic level or number of interactions) is known to affect its ability to persist following disturbance. 
    3. These simple measures, however, offer only a coarse description of how species fit into their community. 
    4. One would therefore expect that more detailed structural measures such as  participation in three-species motifs (meso-scale structures which provide information on a species' direct and indirect interactions) will also be related to probability of persistence.
    5. Disturbances affecting the basal resources have particularly strong effects on the rest of the food web. 
    6. However, how disturbances branch out and affect consumer persistence depends on the structural pattern of species interactions in several steps. 
    7. The magnitude, e.g. proportion of basal resources lost, will likely also affect the outcome.     
    8. Here, we analyse whether a consumer's risk of secondary extinction after removal of basal resources depends on the consumer's motif participation and how this relationship varies with the severity of disturbance.
    9. We show that consumer species which participate more frequently in the direct competition motif and less frequently in the omnivory motif generally have higher probability of persistence following disturbance to basal resources.
    10. However, both the strength of the disturbance and the overall network structure (i.e., connectance) affect the strength and direction of relationships between motif participation and persistence.
    11. Motif participation therefore captures important trends in species persistence and provides a rich description of species' structural roles in their communities, but must be considered in the context of network structure as a whole and of the specific disturbance applied.

@@

    Motifs are unique patterns of $n$ interacting species that correspond to different sets of direct and indirect interactions, providing a holistic description of meso-scale network structure~\citep{Stouffer2007,Stouffer2012}.
    Some motifs are over-represented in stable food webs, suggesting that these motifs provide stability~\citep{Borrelli2015a}. 
    This further suggests that the more frequently species participate in these motifs, the more stable the food web would be.
    % Indeed, dynamic simulations show that participation in some motifs can provide greater persistence after the loss of a single species~\citep{Cirtwill2022Oikos}, suggesting that motif participation may be related to persistence after other types of disturbance.
    The vector of frequencies with which a species appears in each motif --its \emph{motif participation}-- describes the species' position in the meso-scale structure of a food web and, like degree and trophic level, is easily calculated once the structure of the network is known.
    Importantly, motif participation incorporates information on the positions of focal species' interaction partners, as well as potential indirect interactions between species.
    Two species might have the same trophic level and the same number of prey, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists (for an example see Fig.~\ref{fig:concept}).
    Since specialists are usually more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than will a predator of generalists.
    This extra information on the focal species' interaction partners may thus provide extra insight when predicting extinction risk.

    %There are 13 possible three-species motifs (the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}, 
    We define motif participation using the well-studied three-species motifs~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}. 
    Four of these motifs --the three-species chain, apparent competition, direct competition, and omnivory-- are especially interesting as they are by far the most common in empirical networks~\citep{Stouffer2007, giling2019plant} and more stable than other motifs (i.e., are more likely to retain all three species in the motif) when modelled in isolation~\citep{Borrelli2015a}.
    These four motifs also have clear ecological interpretations.
    The three-species chain indicates vertical flows of energy and biomass in a network and captures the possibility of indirect interactions across multiple trophic levels (e.g., between a plant and a predator of the plant's herbivore).
    The dynamics of three-species chains in isolation are well understood from Lotka-Volterra models, and stable equilibria with all three species persisting can often be identified~\citep{Borrelli2015a,Hsu2015}.
    In general, food chains are more likely to be stable when predator-prey body-mass ratios are smaller~\citep{Jonsson1998}.
    The direct competition motif, where two consumers share the same resource, captures the possibility of indirect interactions between consumers by depletion of the shared resource in competitive exclusion~\citep{Hsu2015}. 
   Despite the potential for exclusion where one predator is competitively superior, this motif is highly robust to changes in interaction strength and therefore has high quasi-sign stability~\citep{Borrelli2015a}.
    The apparent competition motif indicates a predator which consumes two resources and, by being somewhat generalist, may be able to persist following the loss or reduction of one prey.
    However, there may also be indirect effects between the linked prey that make the predator more likely to go extinct~\citep{Melian2002}.
    Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes  the potential for a plant to have both direct and indirect effects on the predator. 
    Omnivory is less stable in isolation than the other three motifs~\citep{Borrelli2015a}, though there is a region of stable equilibrium for Lotka-Volterra models of omnivory~\citep{Hsu2015}.
    In whole food webs, frequent omnivory has been proposed to both increase and decrease overall stability in different conditions~\citep{McCann1997,Emmerson2004,Monteiro2016}.
    As most of the above studies either model motifs in isolation or examine overall effects on the whole food web, it remains unclear how a particular species' motif participation will affect its specific extinction risk.

    Previous work using dynamical simulations has shown that participation in some motifs is associated with longer persistence after the loss of a single species in the food web~\citep{Cirtwill2022Oikos}, but the relationship between motif participation and persistence in the face of other disturbances remains unclear.
   As the basal resource community is the foundation upon which all other species depend, and loss of basal resource diversity typically causes declines in both abundance and richness of all types of consumers~\citep{scherber2010bottom,Dobson2009food, Mduma1999food, Georgiadis2007}.
    Moreover, disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant}, and consumer diversity \citep{scherber2010bottom, Baiser2016,li2020bottom}.
    It is therefore crucial to understand how bottom-up effects from disturbance to basal resources affect the rest of the food web.  Here, we are interested in understanding if motif participation can provide additional information on how extinction risks of consumer species are affected by the loss of basal resources, i.e., the effects of bottom-up disturbances.
    As well as identifying patterns of motif participation that are associated with ability to persist after a loss of basal resources, we are particularly interested in whether the strength of  disturbance affects relationships between motif participation and persistence.
    For example, participation in one motif might be associated with greater extinction risk after a mild disturbance but be associated with less risk when when more basal resources are lost if the direct and indirect interactions within the motif amplify the effect of small disturbances while providing `back-up' resources when many prey are removed by a strong disturbance.

@@

        The removal of cycles also affects which motifs can appear in the network. 
        Any motif which contains a cycle is impossible in an acyclic network.
        This means that only four motifs can appear: apparent competition, direct competition, omnivory, and the three-species chain.
        Fortunately, they are also the motifs that are most common in empirical networks~\citep{Stouffer2007}, have been most studied in isolation (e.g.,~\citealp[]{Hastings1991,Holt1987,Bascompte2005,Polis1989,Zabalo2012,Lefevre2009,Holt1997,Kondoh2008,McKinnon2013,Laws2013}), have the clearest ecological interpretations and, because of these three properties, are of greatest interest to us.

        \subsubsection*{Modelling consumers' risk of secondary extinction} %[[trimmed down to here]]

@@

            We also confirmed previously-known relationships between persistence and other measures of network structure or a species' participation in the network.
            To identify the relationship between large-scale network structure and mean persistence within a network, and how the strength of disturbance might affect this relationship, we fit a GLM relating mean persistence to network size, connectance, disturbance, and their interactions.
            As the three-way interaction was significant, we did not simplify the model. 
            To confirm known relationships between persistence and in-degree or trophic level in our simulated networks and identify how they might vary with different strengths of disturbance, we fit two GLMs with binomial error distribution relating a consumer's persistence to in-degree or STL, disturbance, and the interaction between them (\emph{Appendix S3.1}, Eq. S2-3).
            When relating persistence to STL, we initially included a random effect of network size and connectance but removed these terms due to model singularity as above.
            The model for in-degree was non-singular when including these random effects and so they were retained in this case.

@@

        The average persistence of all consumers was lower in networks with higher connectance, did not vary significantly with network size (Table S4, \emph{Appendix S6.1}), and  varied with the frequencies of different motifs in the network's motif profile.
        Specifically, average consumer persistence was higher in networks with higher proportions of the apparent competition motif, lower in networks with higher proportions of the omnivory motif, and was not significantly related to the frequencies of the three-species chain or direct competition motifs (Fig.~\ref{fig:lm_CS}, Table S5, \emph{Appendix S6.2}).
        These relationships are broadly similar to, but weaker than, the relationships between species' probabilities of persistence and motif participation.
        There was no significant interaction between the proportion of any motif in the network motif profile and level of disturbance.

@@

    Disturbance strength is often neglected when predicting risk of extinction, but proved important in our results.
    Several of the relationships we demonstrated between probability of persistence and species roles changed direction when moving from weak to strong disturbance.
    These simulation results echo other warnings about `tipping points' where previously beneficial traits can become risky~\citep{Latty2019,Golubski2016,Tylianakis2014}.
    For example, at low levels of disturbance a high degree may be protective since species can compensate if one or a few prey are lost. 
    At high disturbance, however, a high-degree species is likely to be affected by disturbances through many paths and therefore have lower persistence.
    This context-dependence contrasts with the general assumption that species with more resources will be less vulnerable~\citep{Allesina2012,binzer2011susceptibility}, though high-degree species may be less vulnerable than we expect if they are able to switch to alternative prey.
    Our model does not include interaction rewiring, but it is possible that predators with many observed prey may also be more able to attack new prey species after a disturbance than specialists.
    These disturbance effects mean that efforts to define a set of target species for conservation interactions must take the current (and future) level of disturbance the community experiences into account.

@@

    The authors thank Gy\"{o}rgy Barab\'{a}s for excellent discussion of the ideas in this manuscript. A{\AA} is supported by the Swedish Research Council for Sustainable Development Formas (\#2015-01262) granted to AE. ARC is supported by a Finnish Academy postdoctoral research grant (\#332999). ARC, A\r{A}, and AE conceived the study. ARC and KLW simulated networks. A\r{A} and ARC conducted analyses. All authors contributed to writing the manuscript.

@@

        \begin{figure}[hb!]
        \centering
        % \includegraphics[width=.9\textwidth]{figures/Figure_1.eps}
        \caption{Conceptual figure of a toy food web. \textbf{a)} This food web contains examples of the four motifs: apparent competition (orange), three-species chain (purple), omnivory (red), and direct competition (blue). These motifs capture different ways in which disturbances to basal resources can propagate through the network. For example, a disturbance to the bottom species in a three-species chain will have a direct effect on the middle species but only indirectly affect the top species while a disturbance to the bottom species in an omnivory motif will have a direct effect on the middle species and direct and indirect effects on the top species. 
        Note that the nodes participate in more motifs than the ones highlighted and that a species' motif participation vector is the count of all motifs of each type in which the species appears. \textbf{b)} In the same food web, species with identical degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only three chains while species B (dark grey node with its direct consumption links colored dark grey) participates in seven (see inserted table for counts of participation in all four motifs; species A has the motif participation vector [3, 3, 0, 1] while species B has the motif participation vector [3, 7, 0, 1]).
        Despite having equal degree and trophic level, species A may be at greater risk of going extinct than species B because of its dependence on specialist prey species. Two-thirds of the prey of species B are generalists and therefore more likely to survive a disturbance to basal resources. By including information about indirect interactions, motif profiles reflect this difference between species A and B while trophic level and degree do not.}
    \label{fig:concept}
    \end{figure}

@@

        \begin{figure}[hb!]
        \centering
        % \includegraphics[width=0.95\textwidth]{figures/persistence_motif_participation.eps}
        \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on probability of persistence following disturbance to basal resources (y-axis). The effect of participating in each motif is based on the Equation~\ref{propreq}. The different colored lines indicate the probability of extinction of basal species, from $\pi_{disturbed} = 0.1$ (top, purple; no disturbance) to $\pi_{disturbed} = 0.5$ (bottom, yellow; high disturbance). 95\% confidence intervals for each line are shown in lighter shading. Note that omnivory made up a smaller proportion of species' roles than other motifs; lines are plotted over the observed ranges of motif participation.}
    \label{fig:prop_lmer_all}
    \end{figure}

    \begin{figure}[hb!]
    \centering
        % \includegraphics[width=\textwidth]{figures/Fig4.pdf}
        \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species. Columns show the result for different motifs, rows show results for low, medium, and high connectance, and colour indicates the strength of disturbance to basal resources ($\pi_{disturbed}$; probability of extinction for a basal resource). The dotted, vertical lines indicate zero on the x-axis. A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increases persistence.}
        \label{fig:density_prop}
    \end{figure}

    \begin{figure}[ht!]
        \centering
        % \includegraphics[width=\textwidth]{figures/persistence_vs_SC_lm.eps}
        \caption{The probabilities of persistence for a particular consumer (\textbf{A, B}) and the average probability of persistence across all consumers (\textbf{C-E}) varied with other elements of network structure as well as with motif participation.
        \textbf{A} Persistence of consumers decreased with increasing STL at all levels of disturbance.
        \textbf{B} Persistence of consumers increased with increasing in-degree when the probability of extinction for basal resources was low but decreased with increasing in-degree when the probability of extinction was high.
        \textbf{C,D} Mean persistence of consumers in a network decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance. There was no significant relationship between mean persistence and network size.
        \textbf{E} Mean persistence of consumers decreased with increasing proportions of omnivory in the network's motif profile and increased with increasing proportions of other motifs. Persistence also decreased significantly with increasing probability of basal species extinction but there was no significant interaction; we therefore show trends only for $\pi$=0.1.}
        \label{fig:lm_CS}
    \end{figure}

    \begin{figure}[hb!]
        \centering
        % \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
        \caption{Motif participation correlated with in-degree and trophic level, and both of these simpler role measures were also correlated with persistence. \textbf{A)} The proportion of omnivory increased with increasing degree while all other proportions decreased. \textbf{B)} The proportions of omnivory and direct competition decreased with increasing STL while the proportions of apparent competition and three-species chains increased. \textbf{C-D)} The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition, where the interaction was not significant). To cover the largest range of slopes, we show relationships over connectance for the smallest networks (\textbf{C}) and over network size for the most-connected networks (\textbf{D}). See also Fig. S6.}
        \label{fig:motifs_vs_TL_and_deg}
    \end{figure}
